### 2015-11-11 papers/B1555_EdgeOn/Resubmission/referee_response.txt
Started referee response

We thank the referee for his/her comments about the paper.  We have
responded to the comments below, and modified the text accordingly.
Overall, we have toned down our somewhat sweeping claims.

> Comments to the Author: This paper presents an interesting new
> observational result for the strong gravitational lens B1555+375,
> along with an abundance of poorly supported conjecture.  The
> observations of B1555+375 should be published, but the authors should
> either remove their speculations, or (more usefully) provide evidence
> in support of them.
> 
> CLASS B1555+375 is one of the "poster child" gravitational lens
> systems.  It is famous for its prototypical flux ratio anomaly, which
> has been cited as evidence for dark matter substructure in this
> system.  The new result presented in this paper is near-infrared
> imaging with Keck and HST which reveals a linear structure in the lens
> galaxy emission consistent with an edge-on thin disk.  Most
> intriguingly, this structure passes near the anomalous images, and
> could explain the observed flux anomaly without the need to invoke
> dark matter substructure.  This would be an important result, and
> hence should be published.
> 
> However, the discussion in the paper lacks important quantitative
> details.  For example, the paper does not describe the total
> luminosity of the disky component, compared to the spheroidal
> component of the lens galaxy.  This is important because the lens
> model discussed in the paper requires the disk component to comprise ~
> 10% of the mass within the Einstein radius, and it is unclear if that
> large mass fraction is consistent with the photometry shown in Figure
> 1.
>

RESPONSE: The referee is correct to point out that we did not quantify
the bulge/disk ratio in the lensing galaxy.  We have performed a
bulge+disk fit to the lens-galaxy surface brightness distribution.
Our fit returns a bulge component that is consistent with zero
emission.  [DO WE WANT TO MAKE M/L RATIO CALCULATIONS??].  We have
added text to the paper describing this result.  We note that the two
components in our lens model can be thought of as the edge-on disk and
the dark-matter halo of the lensing galaxy.

> More importantly, the large disk component in the lens model produces
> a significant distortion to the usually ellipsoidal critical curves
> found in strong lenses, as shown in Fig 3.  It has long been known
> that such departures from ellipsoidal density distributions are
> tightly constrained by extended emission from the source galaxies,
> e.g. arcs and Einstein rings that span a large range of azimuthal
> angle.  The imaging shown in Fig 2 exhibits faint extended arclets
> from the source galaxy that could constrain disky components in the
> lens mass distribution, but the lens modeling in the paper neglects to
> fit the extended emission, and instead only models the positions and
> fluxes of the pointlike radio quasar.  The signal-to-noise of the
> extended emission may be too low to provide meaningful constraints,
> but given that the authors have this data in hand, and given that the
> authors have considerable experience modeling this type of extended
> emission in lens systems, they should attempt to fit the observed
> arclets to see if a 10% mass fraction in an off-center thin disk is
> consistent with the observed arclets.

RESPONSE: The referee is correct in pointing out that the modelling of
the extended images would provide tighter constraints on the proposed
model. Unfortunately, in the case of B1555 IR imaging the disk is
responsible for a significant level of dust extinction that leave
images D and B unobservable, and even the arc near image A strongly
affected. We have, however, tried modelling this dataset by only
focusing on image C which is not affected by dust, and the results are
consistent with those found by modelling the point images positions
and fluxes. For the VLBI extended images the poor quality of the data
prevented us from modelling this dataset in a reliable manner.  
[ADD TEXT TO PAPER??]

> 
> The final section of the paper speculates that massive disc components
> may be prevalent in strong lenses, and that such discs may bias
> inferences on the abundance of dark matter substructure using strong
> lensing.  The authors fail to provide evidence supporting these
> speculations, even though they have the data to do so.  As noted
> above, strong lenses with extended source emission (arcs and rings)
> tightly constrain the presence of thin components in lens galaxies.
> In some cases with Einstein rings, disc components are excluded from
> contributing more than percent-level fractions of the mass density.
> It would therefore be highly surprising for a substantial fraction of
> strong lenses to contain hidden, massive edge-on discs that could
> distort the lensing so significantly, as the paper intimates.
> 
> Rather than wildly speculating, the authors could easily check the
> plausibility of this claim.  A number of the authors were members of
> the SLACS collaboration, whose dataset contains many lenses with
> extended emission from the source galaxies.  These systems are the
> obvious way to quantify the fraction of lens systems containing
> massive edge-on discs, since the arcs in many SLACS lenses span a wide
> range of azimuthal angle.  Once they have quantified the abundance of
> massive edge-on discs, the authors should then perform Monte Carlo
> calculations of quasar lenses to calculate the statistics of flux
> ratio anomalies produced by these discs.  Only then can they state
> whether hidden edge-on massive discs can bias observational
> measurements of dark matter substructure with gravitational lens flux
> ratios.

RESPONSE: In the SHARP sample, we find that 5 out of the 25 systems
that we imaged have clear disk components, indicating that this is not
a completely unexpected phenomenon.  While we agree with the referee
that it is important to quantify how common disks are in lensing
galaxies, we do not feel that the SLACS sample provides the
appropriate comparison.  This is because a significant portion of the
SHARP systems come from source-selected samples, such as CLASS, that
are agnostic about the nature of the lensing galaxy.  In contrast,
SLACS lenses were selected to be systems where the lensing galaxy was
known a priori to be a massive elliptical galaxy.  We would therefore
expect a lower incidence of disky systems in SLACS than in the SHARP
sample.  We agree completely with the referee that doing a
quantification of the statistics of massive disks and their effects on
flux-ratio anomalies is an extremely important task and, in fact, this
is one of the future projects that we are planning.  However, we feel
that such an extensive analysis is beyond the scope of this letter.
As such, we have modified the text to focus more on this individual
lens system and put more caveats on the possible generalization of the
results.