### 2015-12-09 papers/B1555_EdgeOn/Resubmission_1208/referee_response.txt
the complete resubmit version

We thank the referee for his/her comments about the paper.  We have
responded to the comments below, and modified the text accordingly.

> Comments to the Author: This paper presents an interesting new
> observational result for the strong gravitational lens B1555+375,
> along with an abundance of poorly supported conjecture.  The
> observations of B1555+375 should be published, but the authors should
> either remove their speculations, or (more usefully) provide evidence
> in support of them.
> 
> CLASS B1555+375 is one of the "poster child" gravitational lens
> systems.  It is famous for its prototypical flux ratio anomaly, which
> has been cited as evidence for dark matter substructure in this
> system.  The new result presented in this paper is near-infrared
> imaging with Keck and HST which reveals a linear structure in the lens
> galaxy emission consistent with an edge-on thin disk.  Most
> intriguingly, this structure passes near the anomalous images, and
> could explain the observed flux anomaly without the need to invoke
> dark matter substructure.  This would be an important result, and
> hence should be published.
> 
> However, the discussion in the paper lacks important quantitative
> details.  For example, the paper does not describe the total
> luminosity of the disky component, compared to the spheroidal
> component of the lens galaxy.  This is important because the lens
> model discussed in the paper requires the disk component to comprise ~
> 10% of the mass within the Einstein radius, and it is unclear if that
> large mass fraction is consistent with the photometry shown in Figure
> 1.
>

RESPONSE: The referee is correct to point out that we did not quantify
the stellar mass in the lensing galaxy.  However, without the
redshifts of the lensing galaxy, we cannot convert the photometry into
a mass estimate.  That being said, the scale radius of the disk is
larger than the Einstein ring radius.  In this regime, studies of
edge-on disk lens galaxies (e.g., Dutton et al. 2011 and 2013) show
that the baryonic component can be significant, often contributing around
40 percent of the circular velocity at 2.2 scale radii.  We have also
performed a bulge+disk fit to the lens-galaxy surface brightness
distribution.  Our fit returns a bulge component that is consistent
with zero emission.  We have added text to the paper describing this
result.  We note that the two components in our lens model can be
thought of as (1) the edge-on disk and (2) the (minimal) bulge +
dark-matter halo of the lensing galaxy.

> More importantly, the large disk component in the lens model produces
> a significant distortion to the usually ellipsoidal critical curves
> found in strong lenses, as shown in Fig 3.  It has long been known
> that such departures from ellipsoidal density distributions are
> tightly constrained by extended emission from the source galaxies,
> e.g. arcs and Einstein rings that span a large range of azimuthal
> angle.  The imaging shown in Fig 2 exhibits faint extended arclets
> from the source galaxy that could constrain disky components in the
> lens mass distribution, but the lens modeling in the paper neglects to
> fit the extended emission, and instead only models the positions and
> fluxes of the pointlike radio quasar.  The signal-to-noise of the
> extended emission may be too low to provide meaningful constraints,
> but given that the authors have this data in hand, and given that the
> authors have considerable experience modeling this type of extended
> emission in lens systems, they should attempt to fit the observed
> arclets to see if a 10% mass fraction in an off-center thin disk is
> consistent with the observed arclets.

RESPONSE: The referee is correct in pointing out that the modelling of
the extended images would provide tighter constraints on the proposed
model. As a matter of fact, we had hoped to do exactly the analysis
that the referee has suggested.  Unfortunately, in the case of B1555
IR imaging the SNR of the lensed extended emission is low.  More
importantly, the disk is responsible for a significant level of dust
extinction that leave images D and B unobservable, and even the arc
near image A strongly affected. We have, however, tried modelling this
dataset by only focusing on image C which is not affected by dust, and
the results are consistent with those found by modelling the point
images positions and fluxes.  Similarly we hoped to use the VLBI
images for the extended source modeling, but as mentioned in the text
the poor quality of the data prevented us from modelling this dataset
in a reliable manner.  We have added text to the manuscript explaining
this.

> 
> The final section of the paper speculates that massive disc components
> may be prevalent in strong lenses, and that such discs may bias
> inferences on the abundance of dark matter substructure using strong
> lensing.  The authors fail to provide evidence supporting these
> speculations, even though they have the data to do so.  As noted
> above, strong lenses with extended source emission (arcs and rings)
> tightly constrain the presence of thin components in lens galaxies.
> In some cases with Einstein rings, disc components are excluded from
> contributing more than percent-level fractions of the mass density.
> It would therefore be highly surprising for a substantial fraction of
> strong lenses to contain hidden, massive edge-on discs that could
> distort the lensing so significantly, as the paper intimates.
> 
> Rather than wildly speculating, the authors could easily check the
> plausibility of this claim.  A number of the authors were members of
> the SLACS collaboration, whose dataset contains many lenses with
> extended emission from the source galaxies.  These systems are the
> obvious way to quantify the fraction of lens systems containing
> massive edge-on discs, since the arcs in many SLACS lenses span a wide
> range of azimuthal angle.  Once they have quantified the abundance of
> massive edge-on discs, the authors should then perform Monte Carlo
> calculations of quasar lenses to calculate the statistics of flux
> ratio anomalies produced by these discs.  Only then can they state
> whether hidden edge-on massive discs can bias observational
> measurements of dark matter substructure with gravitational lens flux
> ratios.

RESPONSE: In the SHARP sample, we find that 5 out of the 25 systems
that we imaged have clear disk components, indicating that this is not
a completely unexpected phenomenon.  While we agree with the referee
that it is important to quantify how common disks are in lensing
galaxies, we do not feel that the SLACS sample provides the
appropriate comparison.  This is because a significant portion of the
SHARP systems come from source-selected samples, such as CLASS, that
are agnostic about the nature of the lensing galaxy.  In fact, CLASS
contains at least one other system, B1600+434 (albeit a double), where
the lensing galaxy is an edge-on disk with a very low bulge/disk
ratio, and at least 2 more disky lenses.  In contrast, SLACS lenses
were selected to be systems where the lensing galaxy was known to be a
luminous red galaxy, and thus likely to be an elliptical.  We would
therefore expect a lower incidence of disc systems in SLACS than in
the SHARP sample.  The SWELLS sample, on the other hand, was selected
to have edge-on disc galaxies as lenses.  Due to the selection effects
in each case, neither of these samples will be much help in
determining the statistics of disc lenses.  The CLASS sample, with 3-4
disk lenses out of 22 systems, might be more informative if it were
not so small.  Overall, we agree completely with the referee that
doing a quantification of the statistics of massive disks and their
effects on flux-ratio anomalies is an extremely important task and, in
fact, this is one of the future projects that we are planning.
However, we feel that such an extensive analysis is beyond the scope
of this letter.  As such, we have modified the text to focus more on
this individual lens system and put more caveats on the possible
generalization of the results.