### 2016-02-01 papers/B1555_EdgeOn/Resubmission_v2/referee_response_2.txt
Added second B1555 referee response to the repo

>The revised version of the paper still contains the unsubstantiated speculation pointed out in my previous report, with no additional supporting evidence.  I am puzzled by the authors' statement in their response that an analysis of the effects of disks on lens statistics would be beyond the scope of the paper, since this is literally the topic of the paper.  The main problem now is that the paper's title is misleading, since nothing in this work provides any reason to believe that there is a bias in observational measurements of dark matter substructure with gravitational lens flux ratios.

RESPONSE: The referee is correct to point out that our title may be misleading in implying that we intend to discuss the effects of disks on lens statistics.  Thus, we add more description on how can we be informed by a new model of a single system in the discussion. That being said, discs with high inclination in lens system can significantly affect the lensing cross-section, which has been shown in Maller et al. 1997, Möller et al. 2003, and Keeton & Kochanek 1998.  Although the fraction of high-inclination spiral lenses may not be high in the general population, the current sample of lensed quasars that have flux-ratio anomalies is fairly small size.  Therefore, if even a few systems within the sample have edge-on discs, these are sufficient to bias the substructure analysis. As the current work of flux anomalous lenses are analyzed with a standard SIE mass profile, it is fair to say that the disc effect is a bias that been ignored for a long time. Moreover, from the high-resolution IR imaging among these flux anomalous lenses, we know that B1555 is not the only system having an edge-on disc across lensed images. Since this is the first time an edge-on disc been confirmed that can be accounted for, at least, part of flux anomalies, we think it is reasonable to keep our title with the wording ‘bias’ rather scope down the implication from our results.

[SOMETHING TO ADD SOMEWHERE ABOVE]:  We seem to have a fundamental disagreement with the referee about the topic of the paper.  The referee feels that it should be an analysis of the statistical effect of disks in lenses.  Most/all previous substructure analyses have interpreted _all_ radio-wavelength flux ratio anomalies as being due to substructure.  We feel that the point of this Letter is to show that the previous assumptions may be wrong, since we have a clear case in which a disc _may_ produce the observed flux-ratio anomaly.  Having made that point in this short letter, we intend to write (and are actively working on) a longer paper that does a full statistical analysis.    
-- We also need to discuss the higher incidence of spirals in small-separation lens systems

>As noted in my previous report, the observational result is quite interesting and merits publication. Since the authors were unable to support their speculations about biases in lens modeling, and have revised the text to focus more on this one specific lens system, I recommend that they amend the title and abstract to more accurately reflect the actual contents of the paper. Something along the lines of "An edge-on disk in the anomalous lens B1555+375" would be more accurate and less misleading than the current title, which is unacceptable for publication given what is actually in the paper.  Similarly, the last 2 sentences of the abstract is misleading.  Although the authors' observations lack sufficient signal/noise to constrain the amplitude of the disk mass in this system, in other lens systems with arcs, the fraction of mass in disky components is constrained to be <~ 1% of the mass in the Einstein radius.  Unless the authors present evidence or arguments for why these previous findings are incorrect, the last sentence of the abstract should be removed.

RESPONSE: We understand that if a constraint from photometry can be applied on B1555+375, it will help us to claim that the best-fit model we found is a ’true’ model. However, the goal of this paper is to demonstrate and bring the attention to the community about the disc-effect. Indeed, as the referee mentioned, the disc mass fraction in the Einstein radius can be down to  <~1%. This is true for massive elliptical lenses but does not hold for spiral lenses especially for lenses having high-inclined discs. The SWELLS samples, 19 spiral lenses, have stellar mass at least >~40% in the Einstein radius. Despise most of SWELLS lenses have a significant bulge therefore they are not perfect analogous systems to B1555+375, it is still unphysical to say that the bulge accounts for more than 97% of stellar mass in the Einstein radius with a clear late-type disc detected.

>Similarly, the penultimate sentence of the abstract is also potentially misleading.  The modeling presented in this paper does not show that simple models containing disks can reproduce the observations, since those models are woefully underconstrained by the data that the authors fit.  Indeed, for the casual reader it would be useful for the text to explicitly state in words that the number of model parameters (13) exceeds the number of observational constraints (11), and that therefore the model should fit the data exactly with 2 exact parameter degeneracies.  Presumably, the weak prior on the centroid locations of the two lens components acts to break the exact degeneracy, though it is puzzling that the uncertainties on the model parameters listed in Table 3 are so much smaller than the 0.05 arcsecond prior on the centroids.  It would be useful for the text to provide an explanation for why the uncertainties on the parameters are so small, since in the absence of the centroid prior there should be 2 exact degeneracies.

RESPONSE: We agree with the referee that the number of free parameters in original model can post a potential threat to our conclusions. We therefore follow the approach in the studies of B1600, another edge-on lens, looking at the photometry and fix the centroid positions, ellipticity, and position angle of the disc component in our best-fit model search. This reduces the number of free parameters from 13 to 9. The predicted image positions and flux ratios in this new model is still agree with the observations, which supports our original statements and conclusions. We also take off the 0.05 arcseoncd prior on the centroids in the best-fit model search. The alignment between two components without using a prior implies that this centroid prior is not necessary.