### 2016-02-23 papers/B1555_EdgeOn/Resubmission_v2/referee_response_2.txt
changes from John [not yet finished]

REFEREE: The revised version of the paper still contains the unsubstantiated speculation pointed out in my previous report, with no additional supporting evidence.  I am puzzled by the authors' statement in their response that an analysis of the effects of disks on lens statistics would be beyond the scope of the paper, since this is literally the topic of the paper.  The main problem now is that the paper's title is misleading, since nothing in this work provides any reason to believe that there is a bias in observational measurements of dark matter substructure with gravitational lens flux ratios.

RESPONSE: The referee is correct to point out that our title may be misleading in implying that we intend to analyze the effects of disks on lens statistics. Thus, we change our title in order to focus on the disc effect itself. Also, we add more text in paper to show how we can be informed by a new model of a single system. That being said, discs with high inclination in lens systems can significantly affect the lensing cross-section, as has been shown in Maller et al. 1997, Möller et al. 2003, and Keeton & Kochanek 1998.  Although the fraction of high-inclination spiral lenses may not be high in the general population, the current sample of lensed quasars that have flux-ratio anomalies has a fairly small size (around 10 in all).  Therefore, if even a few systems within the sample have edge-on discs, these are sufficient to bias the substructure analysis. As previous work on the flux ratios in lensed quasar systems used standard SIE mass profiles, it is fair to say that the disc effect is a bias that has been ignored for a long time. Moreover, from our high-resolution IR imaging of anomalous flux-ratio lenses, we know that B1555 is not the only system that has an edge-on disc across the problematic lensed images. Since this is the first time that an edge-on disc has been detected that can account for at least some part of an observed flux ratio anomaly, we think it is reasonable to talk about this ‘potential bias’ in the discussion.

As described above, we seem to have a disagreement with the referee about the topic of the paper.  The referee feels that it should be a broad analysis of the statistical effect of discs in lenses.  Most/all previous substructure analyses have interpreted _all_ radio-wavelength flux ratio anomalies as being due to substructure.  We feel that the point of this Letter is to show that the previous assumptions may be wrong, since we have a clear case in which a disc _may_ produce the observed flux-ratio anomaly.  Having made that point in this short letter, we intend to write (and are actively working on) a longer paper that does a full statistical analysis.

---------

REFEREE: As noted in my previous report, the observational result is quite interesting and merits publication. Since the authors were unable to support their speculations about biases in lens modeling, and have revised the text to focus more on this one specific lens system, I recommend that they amend the title and abstract to more accurately reflect the actual contents of the paper. Something along the lines of "An edge-on disk in the anomalous lens B1555+375" would be more accurate and less misleading than the current title, which is unacceptable for publication given what is actually in the paper.  Similarly, the last 2 sentences of the abstract is misleading.  Although the authors' observations lack sufficient signal/noise to constrain the amplitude of the disk mass in this system, in other lens systems with arcs, the fraction of mass in disky components is constrained to be <~ 1% of the mass in the Einstein radius.  Unless the authors present evidence or arguments for why these previous findings are incorrect, the last sentence of the abstract should be removed.

RESPONSE: As we mentioned in the previous revision, the fraction of mass within the Einstein ring radius that is due to the disc for the B1555 system is perfectly consistent with the results found in the SWELLS sample of lenses.  That reference in the text was a bit subtle, so we have made this point more explicit in the current version of the paper.  The SWELLS sample is a group of edge-on spiral lenses selected from SDSS spectra.  These systems provide clear evidence of high fractions of stellar mass within the Einstein radius and, furthermore, some of the SWELLS systems also have arcs right across their edge-on discs, similar to what we see in B1555+375.  Below we discuss some specific results from several SWELLS papers that directly address the referee's point.

First, though, we provide a general argument for why finding that the edge-on disc component in B1555 contributes around 15 per cent of the Einstein ring mass is reasonable.  We note that B1555 is a small-separation lens system, with an Einstein ring radius of less than 0.25 arcsec.  It has been known from the early days of lensing (e.g., Turner, Ostriker, and Gott) that spiral lenses should be more common among smaller separation systems, since they tend to have lower mass (and has been seen in observations, e.g. the smallest separation lens B0218+357).  Furthermore, many studies of both local galaxies and more distant lensing galaxies have shown that in the inner portions of galaxies the baryonic component is more important, with the dark matter component becoming increasingly dominant at greater galactocentric distance.  Among these studies, Figure 6 in Brewer et al. 2014 (Paper VI in the SWELLS series) shows that stellar mass dominates the lens mass inside the critical curves and, furthermore, this trend becomes slightly stronger for the smaller-separation systems.  Finally, as mentioned in our first response to the referee, we have fit to the photometry of B1555+375 and found no evidence for a luminous bulge.  Thus, the referee's expectation that <1% of the total mass within the critical curve for this system is in a disc would imply that 99% of the Einstein ring mass is in a dark-matter halo, which is not at all consistent with what is seen in other small-separation lens systems.

All that being said, however, the SWELLS data do provide specific and strong evidence for many edge-on spiral systems in which the disc component makes up more than 1% of the mass within the Einstein ring radius.  By combining the information given in Treu et al. 2011 and Brewer et al. 2014, the disc mass fraction measured within the critical curves of SWELLS lenses ranges from 5% to 30% under a conservative 'minimal' Chabrier ISM assumption. Especially when we look at SWELLS J0820+4847, a good analogy to B1555+375 with an edge-on disc and a clear arc across the disc, we find that it has a 15% disc mass fraction within the Einstein radius. This is very similar to our B1555 model that also has a ~15% disc mass fraction.   Overall, we suspect that the suggested "disc mass fraction <~1% within the Einstein radius" most likely holds for massive elliptical lenses but not for edge-on spiral lenses.

We understand that if a constraint from photometry can be applied on B1555+375, it will help us to put better constraints on our model. However, the goal of this paper is to demonstrate and bring to the attention of the community the effects of edge-on discs.  Please note that we are _not_ claiming that the model presented in this paper is the only correct model.  Far from it.  It is entirely possible that the disc in B1555+375 only contributes part of the flux ratio anomaly.  However, even that statement is important for analyses of the current flux-ratio samples, in which substructure has previously been assumed to be the sole cause of any anomalies.  Our view is that the attribution of at least some portion of the flux-ratio anomaly to a clearly visible disc component is at least as physically motivated as assigning the cause of the anomaly to unseen substructures.

In addition, the suggested title by the referee doesn’t fully encompass the work that is contained in the paper, which is significantly more than the detection of a disc. For example, we present
1) new infrared imaging with Keck and HST that show there is a disc that directly crosses the problem images;
2) new radio VLBA imaging that show image A has a smaller angular size, relative to image B, which demonstrates that there must be a perturbation of the mass model, that is, not due to extinction (seen at optical) or absorption;
3) new radio VLBA imaging that show the radio images are not point-sources as have been previously assumed by Dalal & Kochanek (2002) and others, and so will be less sensitive to substructure, that is, the demagnification will be felt less over the whole arc;
4) new estimates of the flux-ratios using the monitoring data by Koopmans et al. (2003), which was the largest source of uncertainty in the Dalal & Kochanek analysis;
5) a new mass model that includes the disc and can reproduce the observed image positions and flux-ratios without the need for substructure;
6) a potential bias in substructure analysis from flux ratio anomalous quasars due to the small size of the current sample.

Our new title reflects these findings.

------

REFEREE: Similarly, the penultimate sentence of the abstract is also potentially misleading.  The modeling presented in this paper does not show that simple models containing disks can reproduce the observations, since those models are woefully underconstrained by the data that the authors fit.  Indeed, for the casual reader it would be useful for the text to explicitly state in words that the number of model parameters (13) exceeds the number of observational constraints (11), and that therefore the model should fit the data exactly with 2 exact parameter degeneracies.  Presumably, the weak prior on the centroid locations of the two lens components acts to break the exact degeneracy, though it is puzzling that the uncertainties on the model parameters listed in Table 3 are so much smaller than the 0.05 arcsecond prior on the centroids.  It would be useful for the text to provide an explanation for why the uncertainties on the parameters are so small, since in the absence of the centroid prior there should be 2 exact degeneracies.

We agree with the referee that the number of free parameters in original model can post a potential threat to our conclusions. We therefore follow the approach in the studies of B1600 and B1933, two other edge-on disc lenses, looking at the photometry to fix the centroid positions, ellipticity, and position angle of the disc component in our best-fit model search. This reduces the number of free parameters from 13 to 9. This is the typical number of free parameters for a standard SIE plus external shear lens model, with the replacement of an exponential disc component.  The predicted image positions and flux ratios in this new model still agree with the observations, which supports our original statements and conclusions. We also take off the 0.05 arcsecond prior on the centroids in the best-fit model search. The alignment between two components without using a prior implies that this centroid prior is not necessary.