### 2016-08-04 papers/illustris_disk/ms.tex
update illustris disk paper draft

\documentclass[useAMS,usenatbib]{mn2e}
%\documentclass[usenatbib]{mn2e}
\usepackage{graphicx}
%\usepackage[nottoc]{tocbibind}
\setlength{\topmargin}{-1.2cm}

% If your system does not have the AMS fonts version 2.0 installed, then
% remove the useAMS option.
%
% useAMS allows you to obtain upright Greek characters.
% e.g. \umu, \upi etc.  See the section on "Upright Greek characters" in
% this guide for further information.
%
% If you are using AMS 2.0 fonts, bold math letters/symbols are available
% at a larger range of sizes for NFSS release 1 and 2 (using \boldmath or
% preferably \bmath).
%
% The usenatbib command allows the use of Patrick Daly's natbib.sty for
% cross-referencing.
%
% If you wish to typeset the paper in Times font (if you do not have the
% PostScript Type 1 Computer Modern fonts you will need to do this to get
% smoother fonts in a PDF file) then uncomment the next line
% \usepackage{Times}

%%%%% AUTHORS - PLACE YOUR OWN MACROS HERE %%%%%

% Journal definitions
\newcommand{\apj}[1]{ApJ}
\newcommand{\mnras}[1]{MNRAS}
\newcommand{\apjl}[1]{ApJL}
\newcommand{\pasj}[1]{PASJ}
\newcommand{\aj}[1]{AJ}
\newcommand{\nat}[1]{Nature}
\newcommand{\aap}{A\&A}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title[Flux ratio anomalies from disks]{
%Gravitational Lens Modelling of flux-ratio Anomalous System B1555+375
SHARP IV: Flux ratio anomalies from disks in the Illustris simulation
}
\author[Hsueh et al.]{Jen-Wei Hsueh$^{1}$\thanks{E-mail:
jwhsueh@ucdavis.edu (JWH); otheremail@otheraddress (ANO)} and A. N.
Other$^{2}$\\
$^{1}$Physics Dept., University of California, Davis, 1 Shields Ave.
Davis, CA 95616, USA\\
$^{2}$Building, Institute, Street Address, City, Code, Country}
\begin{document}

%\date{Accepted 1988 December 15. Received 1988 December 14; in original form 1988 October 11}

\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2015}

\maketitle

\label{firstpage}

\begin{abstract}

\end{abstract}

\begin{keywords}
gravitational lensing
\end{keywords}

\section{Introduction}
\section{The Illustris simulation}
\section{Disks selection}
\begin{itemize}
\item discuss how realistic disks in illustris are
\end{itemize}
\subsection{Morphological selection}
\subsection{Kinematical selection I}
\subsection{Kinematical selection II}
\section{Disk properties}
\begin{itemize}
\item how many galaxies have disks
\item mass in disk
\item redshift evolution
\item comparison with B1555 and SWELLS
\item observability of the disks
\end{itemize}
\section{Flux ratio anomalies from disks}

\section{Discussion}
\section*{Acknowledgments}

\bibliographystyle{mn2e}
\bibliography{reference.bib}

\label{lastpage}

\end{document}