### 2016-08-11 papers/illustris_disk/ms.tex
update illustris paper draft

\documentclass[useAMS,usenatbib]{mn2e}
%\documentclass[usenatbib]{mn2e}
\usepackage{graphicx}
%\usepackage[nottoc]{tocbibind}
\setlength{\topmargin}{-1.2cm}

% If your system does not have the AMS fonts version 2.0 installed, then
% remove the useAMS option.
%
% useAMS allows you to obtain upright Greek characters.
% e.g. \umu, \upi etc.  See the section on "Upright Greek characters" in
% this guide for further information.
%
% If you are using AMS 2.0 fonts, bold math letters/symbols are available
% at a larger range of sizes for NFSS release 1 and 2 (using \boldmath or
% preferably \bmath).
%
% The usenatbib command allows the use of Patrick Daly's natbib.sty for
% cross-referencing.
%
% If you wish to typeset the paper in Times font (if you do not have the
% PostScript Type 1 Computer Modern fonts you will need to do this to get
% smoother fonts in a PDF file) then uncomment the next line
% \usepackage{Times}

%%%%% AUTHORS - PLACE YOUR OWN MACROS HERE %%%%%

% Journal definitions
\newcommand{\apj}[1]{ApJ}
\newcommand{\mnras}[1]{MNRAS}
\newcommand{\apjl}[1]{ApJL}
\newcommand{\pasj}[1]{PASJ}
\newcommand{\aj}[1]{AJ}
\newcommand{\nat}[1]{Nature}
\newcommand{\aap}{A\&A}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title[Flux ratio anomalies from disks]{
%Gravitational Lens Modelling of flux-ratio Anomalous System B1555+375
SHARP IV: Flux ratio anomalies from disks in the Illustris simulation
}
\author[Hsueh et al.]{Jen-Wei Hsueh$^{1}$\thanks{E-mail:
jwhsueh@ucdavis.edu (JWH); otheremail@otheraddress (ANO)} and A. N.
Other$^{2}$\\
$^{1}$Physics Dept., University of California, Davis, 1 Shields Ave.
Davis, CA 95616, USA\\
$^{2}$Building, Institute, Street Address, City, Code, Country}
\begin{document}

%\date{Accepted 1988 December 15. Received 1988 December 14; in original form 1988 October 11}

\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2015}

\maketitle

\label{firstpage}

\begin{abstract}

\end{abstract}

\begin{keywords}
gravitational lensing
\end{keywords}

\section{Introduction}
\section{The Illustris simulation}
\section{Disks selection}
\begin{itemize}
\item discuss how realistic disks in illustris are
\end{itemize}
Resolution: stellar mass cut between 7E9Msun~1E10Msun. Due to the resolution limitation, our disk galaxy samples are from the high-mass range of disk galaxy population, with mass range [].

\subsection{Morphological selection}
Photometric properties are commonly used in surveys to identify the morphology type of galaxies. Star particles in simulation can also be visualized (SUNRISE) and provide photometric observants for individual galaxy. We apply the following two criteria to identity a disk galaxy in Illustris simulation: 1) The galaxy has a 
better fit (i.e. a smaller $\chi^2$) in exponential disk light profile than in de Vaucouleur light profile, and 2) the galaxy has a Sersic index smaller than two. The sample selected by these photometric properties are called morphological selection in this work.

\subsection{Kinematical selection I}
In kinematical view, stellar disk is a group of stars co-rotate in the same direction around the  symmetric axis (we will call this symmetric axis ``z-axis,'' for short.) One of the advantages in simulations is that one can access to the full detail of kinematical properties for single mass particle. These kinematical properties provide an approach to decompose different components in disk galaxy: thin disk, thick disk, and spheroid. The dynamical decomposition can be done in the parameter space of the z-component of specific angular momemtum, $J_z$, and the specific binding energy, $E$, of each star (Abadi, et al. 2003). On the $J_z - E$ plane, stellar and gaseous components co-rorating in the same direction define a solid curve $J_{circ}(E)$ (see Fig.2 in Abadi, et al. 2003), which represents the coratating circular orbits in the disk as a function of specific binding energy. The circularity parameter $\epsilon_z$ of each star is defined as the ratio between its  $J_z$ and corresponding $J_{circ}(E)$, showing how closely a stellar or gaseous component follows the circular motion of the disk. In general, the gaseous disk and the thin disk stellar components have a sharp distribution of $\epsilon_z$ peaked about unity. Base on the $\epsilon_z$ distribution, all stars within the ``luminous'' radius can be decomposed into three different dynamical components: a spheriod (including bulge and halo stars), thin disk (stars with $\epsilon_z > 0.7$), and thick disk. Galaxies with more than 40 per cent of stars in thin disk are considered as late-type spirals.

\subsection{Kinematical selection II}
Having more than 40 per cent of stars in thin disk is a strict criteria which selects the galaxies with a large disk and a small bulge. However, spirals with a smaller thin disk but large enough thick disk may also provide sufficient gravitational perturbation to influence lens image flux ratios. Thus we select galaxies having more than 40 per cent stars in their thin disk and thick disk, in another word, galaxies having less than 60 per cent of stars in bulge and halo as the second kinematical selection of disk galaxies.

\section{Disk properties}
\begin{itemize}
\item how many galaxies have disks
\item mass in disk
\item redshift evolution
\item comparison with B1555 and SWELLS
\item observability of the disks
\end{itemize}
\section{Flux ratio anomalies from disks}

\section{Discussion}
\section*{Acknowledgments}

\bibliographystyle{mn2e}
\bibliography{reference.bib}

\label{lastpage}

\end{document}